---
title: "nflData2022"
date: "2023-09-27"
author: "Eric Chu"
output:
  html_document:
    toc: yes            # creates table of contents
    toc_depth: 4        # toc will include headers <= ####
    toc_float: yes      # toc always on left of page
    code_folding: show  # allows hiding of code
---

The following code is used to generate the data plots.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(zoom)
library(plotrix)
```

## Importing CSVs
``` {r Importing all NFL Data}
nfl2022 <- read.csv("nfl2022.csv", header = TRUE)
driveAvg <- read.csv("driveAvg2022.csv", header = TRUE)
passOff <- read.csv("passOff2022.csv", header = TRUE)
rushOff <- read.csv("rushOff2022.csv", header = TRUE)
scoreOff <- read.csv("scoreOff2022.csv", header = TRUE)

"I added the abbrevation column to each dataframe, this will make plotting, sorting, and organization easier."
```


```{r Sort DFs Alphabetically}
nfl2022 <- nfl2022[order(nfl2022$Tm, decreasing = FALSE), ]
driveAvg <- driveAvg[order(driveAvg$Tm, decreasing = FALSE), ]
passOff <- passOff[order(passOff$Tm, decreasing = FALSE), ]
rushOff <- rushOff[order(rushOff$Tm, decreasing = FALSE), ]
scoreOff <- scoreOff[order(scoreOff$Tm, decreasing = FALSE), ]

"In order to have the stats properly correspond to each team, I decided to use alphabetical order since it is the most consistent ordering between each data frame."
```

``` {r Allow viewer to see head for each data frame}
head(nfl2022)
head(driveAvg)
head(passOff)
head(rushOff)
head(scoreOff)
```


## Passing Success and Offensive Success
``` {r Plotting Passing Graphs}

# Passing Attempts vs Offensive Rank
plot(passOff$Att, scoreOff$Pts.G, type = "n", xlab = "Num. of Passing Atts.", ylab = "Points per Game")
text(passOff$Att, scoreOff$Pts.G, labels = passOff$Abv.)
cor(passOff$Att, scoreOff$Pts.G)

# Passing TDs vs Offensive Rank
plot(passOff$TD, scoreOff$Pts.G, type = "n", xlab = "Num. of Passing TDs", ylab = "Points per Game")
text(passOff$TD, scoreOff$Pts.G, labels = passOff$Abv)
cor(passOff$TD, scoreOff$Pts.G)

# Passing Yards vs Offensive Rank
plot(passOff$Yds, scoreOff$Pts.G, type = "n", xlab = "Num. of Passing YDs", ylab = "Points per Game")
text(passOff$Yds, scoreOff$Pts.G, labels = passOff$Abv)
cor(passOff$Yds, scoreOff$Pts.G)

# Passing Y/A vs Offensive Rank
plot(passOff$Y.A, scoreOff$Pts.G, type = "n", xlab = "Passing Yards per Attempt", ylab = "Points per Game")
text(passOff$Y.A, scoreOff$Pts.G, labels = passOff$Abv)
cor(passOff$Y.A, scoreOff$Pts.G)

# Passer Rating vs Offensive Rank
plot(passOff$Rate, scoreOff$Pts.G, type = "n", xlab = "Passer Rating", ylab = "Points per Game")
text(passOff$Rate, scoreOff$Pts.G, labels = passOff$Abv)
cor(passOff$Rate, scoreOff$Pts.G)

```

## Rushing Success and Offensive Success
```{r Plotting Rushing Graphs}
# Rushing Attempts vs Offensive Rank
plot(rushOff$Att, scoreOff$Pts.G, type = "n", xlab = "Num. of Rushing Atts.", ylab = "Points per Game")
text(rushOff$Att, scoreOff$Pts.G, labels = rushOff$Abv.)
cor(rushOff$Att, scoreOff$Pts.G)

# Rushing TDs vs Offensive Rank
plot(rushOff$TD, scoreOff$Pts.G, type = "n", xlab = "Num. of Rushing TDs", ylab = "Points per Game")
text(rushOff$TD, scoreOff$Pts.G, labels = rushOff$Abv)
cor(rushOff$TD, scoreOff$Pts.G)

# Rushing YDs vs Offensive Rank
plot(rushOff$Yds, scoreOff$Pts.G, type = "n", xlab = "Num. of Rushing Yds", ylab = "Points per Game")
text(rushOff$Yds, scoreOff$Pts.G, labels = rushOff$Abv)
cor(rushOff$Yds, scoreOff$Pts.G)

# Rushing Y/A vs Offensive Rank
plot(rushOff$Y.A, scoreOff$Pts.G, type = "n", xlab = "Rushing Y/A", ylab = "Points per Game")
text(rushOff$Y.A, scoreOff$Pts.G, labels = rushOff$Abv.)
cor(rushOff$Y.A, scoreOff$Pts.G)
```

## Passing Success and Team Success
```{r Plotting Passing Win Percentage Graphs}

#Pass Attempts vs WL%
plot(passOff$Att, nfl2022$W.L., type = "n", xlab = "Num. of Pass Atts.", ylab = "Win-Loss Percentage")
text(passOff$Att, nfl2022$W.L., labels = passOff$Abv.)

'
plot(passOff$Att[passOff$Att < 600 & passOff$Att > 550 & nfl2022$W.L.< 0.55 & nfl2022$W.L. > 0.5], nfl2022$W.L.[passOff$Att < 600 & passOff$Att > 550 & nfl2022$W.L.< 0.55 & nfl2022$W.L. > 0.5]
, type = "n", xlab = "Pass Attempts", ylab = "Win-Loss Percentage")
text(passOff$Att, nfl2022$W.L., labels = passOff$Abv.)
'

cor(passOff$Att, nfl2022$W.L.)

# Passing TDs vs WL%
plot(passOff$TD, nfl2022$W.L., type = "n", xlab = "Num. of Passing TDs", ylab = "Win-Loss Percentage")
text(passOff$TD, nfl2022$W.L., labels = passOff$Abv)
cor(passOff$TD, nfl2022$W.L.)

# Passing Yards vs  WL%
plot(passOff$Yds, nfl2022$W.L., type = "n", xlab = "Num. of Passing Yds", ylab = "Win-Loss Percentage")
text(passOff$Yds, nfl2022$W.L., labels = passOff$Abv)
cor(passOff$Yds, nfl2022$W.L.)

# Passing Y/A vs  WL%
plot(passOff$Y.A, nfl2022$W.L., type = "n", xlab = "Passing Y/A", ylab = "Win-Loss Percentage")
text(passOff$Y.A, nfl2022$W.L., labels = passOff$Abv)
cor(passOff$Y.A, nfl2022$W.L.)

# Passer Rating vs  WL%
plot(passOff$Rate, nfl2022$W.L., type = "n", xlab = "Passer Rating", ylab = "Win-Loss Percentage")
text(passOff$Rate,nfl2022$W.L., labels = passOff$Abv)
cor(passOff$Rate, nfl2022$W.L.)
```

## Rushing Success and Team Success
```{r Plotting Rushing Win Percentage Graphs}
# Rushing Attempts vs  WL%
plot(rushOff$Att, nfl2022$W.L., type = "n", xlab = "Num. of Rushing Atts.", ylab = "Win-Loss Percentage")
text(rushOff$Att, nfl2022$W.L., labels = rushOff$Abv.)
cor(rushOff$Att, nfl2022$W.L.)

# Rushing TDs vs  WL%
plot(rushOff$TD, nfl2022$W.L., type = "n", xlab = "Num. of Rushing TDs", ylab = "Win-Loss Percentage")
text(rushOff$TD, nfl2022$W.L., labels = rushOff$Abv)
cor(rushOff$TD, nfl2022$W.L.)

# Rushing YDs vs  WL%
plot(rushOff$Yds, nfl2022$W.L., type = "n", xlab = "Num. of Rushing Yds", ylab = "Win-Loss Percentage")
text(rushOff$Yds, nfl2022$W.L., labels = rushOff$Abv)
cor(rushOff$Yds, nfl2022$W.L.)

# Rushing Y/A vs  WL%
plot(rushOff$Y.A, nfl2022$W.L., type = "n", xlab = "Rushing Y/A", ylab = "Win-Loss Percentage")
text(rushOff$Y.A, nfl2022$W.L., labels = rushOff$Abv.)
cor(rushOff$Y.A, nfl2022$W.L.)
```

## Consider the Balance of Passing vs Rushing
```{r Creating TD and Yardage Proportions Between Passing and Rushing Stats}
'Note: The total TDs are strictly from the offense, so pick-sixes, punt return TDs, and so on will not count'

RushTDsProp <- rushOff$TD/(rushOff$TD + passOff$TD)
RushTDsProp <- cbind(rushOff$Abv., RushTDsProp)

PassTDsProp <- passOff$TD/(rushOff$TD + passOff$TD)
PassTDsProp <- cbind(passOff$Abv., PassTDsProp)

RushYdsProp <- rushOff$Yds/(rushOff$Yds + passOff$Yds)
RushYdsProp <- cbind(rushOff$Abv., RushYdsProp)

PassYdsProp <- passOff$Yds/(rushOff$Yds + passOff$Yds)
PassYdsProp <- cbind(passOff$Abv., PassYdsProp)

"I'm sure there are better ways to combine the prortion of yards and TDs with team abbreviations, but this will do for now."
```

```{r Print New Dataframes}
#Show head of all the new data frames
print(RushTDsProp)
print(PassTDsProp)
print(RushYdsProp)
print(PassYdsProp)
```


## Offensive Success Based on Passing and Rushing Proportions
```{r Plotting Yardage and TD Proprtions vs Offensive Rank}

#Passing Yds Prop vs Offensive Rank
plot(PassYdsProp[,2], scoreOff$Pts.G, type = "n", xlab = "Passing Yds Percentage", ylab = "Points per Game")
text(PassYdsProp[,2], scoreOff$Pts.G, labels = passOff$Abv.)
cor(as.numeric(PassYdsProp[,2]), scoreOff$Pts.G)

#Passing TDs Prop vs Offensive Rank
plot(PassTDsProp[,2], scoreOff$Pts.G, type = "n", xlab = "Passing TDs Percentage", ylab = "Points per Game")
text(PassTDsProp[,2], scoreOff$Pts.G, labels = passOff$Abv.)
cor(as.numeric(PassTDsProp[,2]), scoreOff$Pts.G)

#Rushing Yds Prop vs Offensive Rank
plot(RushYdsProp[,2], scoreOff$Pts.G, type = "n", xlab = "Rushing Yds Percentage", ylab = "Points per Game")
text(RushYdsProp[,2], scoreOff$Pts.G, labels = passOff$Abv.)
cor(as.numeric(RushYdsProp[,2]), scoreOff$Pts.G)

# Rushing TDs Prop vs Offensive Rank
plot(RushTDsProp[,2], scoreOff$Pts.G, type = "n", xlab = "Rushing TDs Percentage", ylab = "Points per Game")
text(RushTDsProp[,2], scoreOff$Pts.G, labels = passOff$Abv.)
cor(as.numeric(RushTDsProp[,2]), scoreOff$Pts.G)

"Because cbind typecasted the proportions to characters (due to the fact that we binded the values with abv. column which is a character vector) 
I had to use the as.numeric function"
```

## Team Success Based on Passing and Rushing Proportions
```{r Plotting Yardage and TD Proprtions vs W/L Percentage}
#Pasing Yds Prop vs Offensive Rank
plot(PassYdsProp[,2], nfl2022$W.L., type = "n", xlab = "Passing Yds Proportion", ylab = "Win/Loss Percentage")
text(PassYdsProp[,2], nfl2022$W.L., labels = passOff$Abv.)
cor(as.numeric(PassYdsProp[,2]), nfl2022$W.L.)

#Passing TDs Prop vs Offensive Rank
plot(PassTDsProp[,2], nfl2022$W.L., type = "n", xlab = "Passing TDs Proportion", ylab = "Win/Loss Percentage")
text(PassTDsProp[,2], nfl2022$W.L., labels = passOff$Abv.)
cor(as.numeric(PassTDsProp[,2]),  nfl2022$W.L.)

#Rushing Yds Prop vs Offensive Rank
plot(RushYdsProp[,2], nfl2022$W.L., type = "n", xlab = "Rushing TDs Proportion", ylab = "Win/Loss Percentage")
text(RushYdsProp[,2],  nfl2022$W.L., labels = passOff$Abv.)
cor(as.numeric(RushYdsProp[,2]),  nfl2022$W.L.)

# Rushing TDs Prop vs Offensive Rank
plot(RushTDsProp[,2], nfl2022$W.L., type = "n", xlab = "Rushing Yds Proportion", ylab = "Win/Loss Percentage")
text(RushTDsProp[,2],  nfl2022$W.L., labels = passOff$Abv.)
cor(as.numeric(RushTDsProp[,2]),  nfl2022$W.L.)
```

## Other Calculations
```{r}
sd(rushOff$Yds)
sd(passOff$Yds)
sd(rushOff$TD)
sd(passOff$TD)
sd(passOff$Rate)
```

## Passing Offense vs Rushing Y/A
```{r}
plot(rushOff$Y.A, passOff$Rate, type = "n", xlab = "Rushing Y/A", ylab = "Passer Rating")
text(rushOff$Y.A, passOff$Rate, labels = rushOff$Abv.)
cor(rushOff$Y.A, passOff$Rate)

plot(rushOff$Yds, passOff$Rate, type = "n", xlab = "Rushing Yards", ylab = "Passer Rating")
text(rushOff$Yds, passOff$Rate, labels = rushOff$Abv.)
cor(rushOff$Yds, passOff$Rate)

plot(rushOff$Y.A, passOff$Yds, type = "n", xlab = "Rushing Y/A", ylab = "Passing Yards")
text(rushOff$Y.A, passOff$Yds, labels = rushOff$Abv.)
cor(rushOff$Y.A, passOff$Yds)

plot(rushOff$Yds, passOff$Yds, type = "n", xlab = "Rushing Yards", ylab = "Passing Yards")
text(rushOff$Yds, passOff$Yds, labels = rushOff$Abv.)
cor(rushOff$Yds, passOff$Yds)

plot(passOff$Yds, rushOff$TD, type = "n", xlab = "Passing Yards", ylab = "Rushing TDs")
text(passOff$Yds, rushOff$TD, labels = rushOff$Abv.)
cor(passOff$Yds, rushOff$TD)


```


